quollr: An R Package for Visalizing 2D Models in High Dimensional Space

### 2024-02-22 paper-quollr.Rmd
Add the hexagonal binning version 2 which initialize the parameters x_start, y_start, buffer_x, buffer_y, hex_size. Now need to find limitations for each parameters and recheck the default for hex_size

```{r}
all_centroids_df <- generate_full_grid_centroids(nldr_df = s_curve_noise_umap, 
                                                 x = "UMAP1", y = "UMAP2", 
                                                 num_bins_x = num_bins_x, 
                                                 num_bins_y = num_bins_y, 
                                                 buffer_x = NA, buffer_y = NA, 
                                                 hex_size = 0.2)

@@

```{r}
ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_point(data = all_centroids_df, aes(x = x, y = y), color = "red") +
  coord_fixed()
```

3. Map hexagonal IDs

- Filter the data set with specific y value

- Order the x values for a specific y value

- Repeat the process for all unique y values

```{r}
full_grid_with_hexbin_id <- map_hexbin_id(all_centroids_df)

ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_text(data = full_grid_with_hexbin_id, aes(x = c_x, y = c_y, label = hexID)) +
  coord_fixed()
```

4. Map polygon IDs

- Filter specific hexagon

- Filter specific polygon

- Check the selected hexagonal centroid exists within the polygon

- if so assign that id to centroid, if not check until find the polygon which contains the centroid

```{r}
full_grid_with_polygon_id <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
```

@@

```{r}
s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, full_grid_with_hexbin_id)
```

@@

```{r}
hex_full_count_df <- generate_full_grid_info(full_grid_with_polygon_id, df_with_std_counts, hex_grid)
```

@@

```{r}
ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
  coord_fixed()

>>>>

```{r}
all_centroids_df <- generate_full_grid_centroids(nldr_df = s_curve_noise_umap, 
                                                 x = "UMAP1", y = "UMAP2", 
                                                 num_bins_x = num_bins_x, 
                                                 num_bins_y = num_bins_y, 
                                                 x_start = NA, y_start = NA, 
                                                 buffer_x = NA, buffer_y = NA, 
                                                 hex_size = 0.2)

@@

```{r}
ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_point(aes(x = c_x, y = c_y), color = "red") +
  coord_fixed()
```

```{r}
full_grid_centroids_with_hexbin_id <- hex_grid |>
  dplyr::select("c_x", "c_y", "hexID") |>
    dplyr::distinct()

ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
  coord_fixed()
```

@@

```{r}
s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, full_grid_centroids_with_hexbin_id)
```

@@

```{r}
hex_full_count_df <- generate_full_grid_info(hex_grid, df_with_std_counts)
```

@@

```{r}
ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = id, fill = std_counts)) +
  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
  coord_fixed()

### 2024-02-22 paper-quollr.tex
Add the hexagonal binning version 2 which initialize the parameters x_start, y_start, buffer_x, buffer_y, hex_size. Now need to find limitations for each parameters and recheck the default for hex_size

\begin{verbatim}
#> [1] 3
\end{verbatim}

@@

\begin{verbatim}
#> [1] 6
\end{verbatim}

@@

\begin{verbatim}
all_centroids_df <- generate_full_grid_centroids(nldr_df = s_curve_noise_umap, 
                                                 x = "UMAP1", y = "UMAP2", 
                                                 num_bins_x = num_bins_x, 
                                                 num_bins_y = num_bins_y, 
                                                 buffer_x = NA, buffer_y = NA, 
                                                 hex_size = 0.2)

@@

\begin{verbatim}
#> Rows: 18
#> Columns: 2
#> $ x <dbl> 0.1464102, 0.4928203, 0.8392305, 0.3196152, 0.6660254, 1.0124356, 0.~
#> $ y <dbl> 0.1845299, 0.1845299, 0.1845299, 0.4845299, 0.4845299, 0.4845299, 0.~
\end{verbatim}

@@

\begin{verbatim}
#> Rows: 108
#> Columns: 3
#> $ x  <dbl> 0.31961524, 0.31961524, 0.14641016, -0.02679492, -0.02679492, 0.146~
#> $ y  <dbl> 0.28412287, 0.08493702, -0.01465590, 0.08493702, 0.28412287, 0.3837~
#> $ id <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4, 4~
\end{verbatim}

\begin{verbatim}
ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_point(data = all_centroids_df, aes(x = x, y = y), color = "red") +
  coord_fixed()
\end{verbatim}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Map hexagonal IDs
\end{enumerate}

\begin{itemize}
\item
  Filter the data set with specific y value
\item
  Order the x values for a specific y value
\item
  Repeat the process for all unique y values
\end{itemize}

\begin{verbatim}
full_grid_with_hexbin_id <- map_hexbin_id(all_centroids_df)

ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_text(data = full_grid_with_hexbin_id, aes(x = c_x, y = c_y, label = hexID)) +
  coord_fixed()
\end{verbatim}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Map polygon IDs
\end{enumerate}

\begin{itemize}
\item
  Filter specific hexagon
\item
  Filter specific polygon
\item
  Check the selected hexagonal centroid exists within the polygon
\item
  if so assign that id to centroid, if not check until find the polygon which contains the centroid
\end{itemize}

\begin{verbatim}
full_grid_with_polygon_id <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
\end{verbatim}

@@

\begin{verbatim}
s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, full_grid_with_hexbin_id)
\end{verbatim}

@@

\begin{verbatim}
hex_full_count_df <- generate_full_grid_info(full_grid_with_polygon_id, df_with_std_counts, hex_grid)
\end{verbatim}

@@

@@

\begin{verbatim}
ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
  coord_fixed()
\end{verbatim}

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-17-1.pdf}

@@

\begin{verbatim}
#>            x         y hexID std_counts
#> 1  0.1464102 0.1845299     1 1.00000000
#> 2  0.4928203 0.1845299     2 0.04761905
#> 3  0.3196152 0.4845299     4 0.19047619
#> 4  0.6660254 0.4845299     5 0.23809524
#> 5  0.4928203 0.7845299     8 0.19047619
#> 6  0.8392305 0.7845299     9 0.33333333
#> 7  0.6660254 1.0845299    11 0.28571429
#> 8  1.0124356 1.0845299    12 0.04761905
#> 9  0.8392305 1.3845299    15 0.38095238
#> 10 0.6660254 1.6845299    17 0.47619048
#> 11 1.0124356 1.6845299    18 0.38095238
\end{verbatim}

@@

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-21-1.pdf}

@@

\begin{verbatim}
#> Rows: 75
#> Columns: 4
#> $ pred_UMAP_1 <dbl> 0.1464102, 0.8392305, 1.0124356, 0.1464102, 0.3196152, 0.6~
#> $ pred_UMAP_2 <dbl> 0.1845299, 0.7845299, 1.0845299, 0.1845299, 0.4845299, 1.6~
#> $ ID          <int> 1, 2, 3, 4, 6, 7, 8, 9, 11, 12, 14, 15, 16, 17, 19, 20, 21~
#> $ pred_hb_id  <int> 1, 9, 12, 1, 4, 17, 9, 5, 9, 18, 5, 1, 18, 1, 5, 15, 2, 15~
\end{verbatim}

@@

\begin{verbatim}
#> $mse
#> [1] 0.372034
#> 
#> $aic
#> [1] -365.1042
\end{verbatim}

@@

@@

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-28-1.pdf}

>>>>

\begin{verbatim}
#> [1] 4
\end{verbatim}

@@

\begin{verbatim}
#> [1] 7
\end{verbatim}

@@

\begin{verbatim}
all_centroids_df <- generate_full_grid_centroids(nldr_df = s_curve_noise_umap, 
                                                 x = "UMAP1", y = "UMAP2", 
                                                 num_bins_x = num_bins_x, 
                                                 num_bins_y = num_bins_y, 
                                                 x_start = NA, y_start = NA, 
                                                 buffer_x = NA, buffer_y = NA, 
                                                 hex_size = 0.2)

@@

\begin{verbatim}
#> Rows: 28
#> Columns: 2
#> $ x <dbl> 0.0000000, 0.3464102, 0.6928203, 1.0392305, 0.1732051, 0.5196152, 0.~
#> $ y <dbl> 0.0, 0.0, 0.0, 0.0, 0.3, 0.3, 0.3, 0.3, 0.6, 0.6, 0.6, 0.6, 0.9, 0.9~
\end{verbatim}

@@

\begin{verbatim}
#> Rows: 168
#> Columns: 6
#> $ c_x   <dbl> 0.0000000, 0.0000000, 0.0000000, 0.0000000, 0.0000000, 0.0000000~
#> $ c_y   <dbl> 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0,~
#> $ x     <dbl> 0.1732051, 0.1732051, 0.0000000, -0.1732051, -0.1732051, 0.00000~
#> $ y     <dbl> 0.09959292, -0.09959292, -0.19918584, -0.09959292, 0.09959292, 0~
#> $ id    <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4~
#> $ hexID <int> 1, 1, 1, 1, 1, 1, 2, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4~
\end{verbatim}

\begin{verbatim}
ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_point(aes(x = c_x, y = c_y), color = "red") +
  coord_fixed()
\end{verbatim}

@@

\begin{verbatim}
full_grid_centroids_with_hexbin_id <- hex_grid |>
  dplyr::select("c_x", "c_y", "hexID") |>
    dplyr::distinct()

ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
  coord_fixed()
\end{verbatim}

@@

@@

\begin{verbatim}
s_curve_noise_umap_with_id <- assign_data(s_curve_noise_umap, full_grid_centroids_with_hexbin_id)
\end{verbatim}

@@

\begin{verbatim}
hex_full_count_df <- generate_full_grid_info(hex_grid, df_with_std_counts)
\end{verbatim}

@@

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-15-1.pdf}

\begin{verbatim}
ggplot(data = hex_full_count_df, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = id, fill = std_counts)) +
  geom_text(aes(x = c_x, y = c_y, label = hexID)) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff") +
  coord_fixed()
\end{verbatim}

@@

@@

\begin{verbatim}
#>            x   y hexID std_counts
#> 1  0.0000000 0.0     1 0.40000000
#> 2  0.3464102 0.0     2 0.26666667
#> 3  0.1732051 0.3     5 0.86666667
#> 4  0.5196152 0.3     6 0.13333333
#> 5  0.3464102 0.6    10 0.20000000
#> 6  0.6928203 0.6    11 0.53333333
#> 7  0.8660254 0.9    15 0.80000000
#> 8  0.6928203 1.2    19 0.06666667
#> 9  0.5196152 1.5    22 0.06666667
#> 10 0.8660254 1.5    23 1.00000000
#> 11 0.6928203 1.8    27 0.46666667
#> 12 1.0392305 1.8    28 0.20000000
\end{verbatim}

@@

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-20-1.pdf}

@@

\begin{verbatim}
#> Rows: 75
#> Columns: 4
#> $ pred_UMAP_1 <dbl> 0.1732051, 0.8660254, 0.8660254, 0.0000000, 0.5196152, 0.6~
#> $ pred_UMAP_2 <dbl> 0.3, 0.9, 0.9, 0.0, 0.3, 1.8, 0.9, 0.6, 0.9, 1.8, 0.6, 0.3~
#> $ ID          <int> 1, 2, 3, 4, 6, 7, 8, 9, 11, 12, 14, 15, 16, 17, 19, 20, 21~
#> $ pred_hb_id  <int> 5, 15, 15, 1, 6, 27, 15, 11, 15, 28, 11, 5, 23, 5, 11, 22,~
\end{verbatim}

@@

\begin{verbatim}
#> $mse
#> [1] 0.2612147
#> 
#> $aic
#> [1] -536.7666
\end{verbatim}

@@

\includegraphics{paper-quollr_files/figure-latex/unnamed-chunk-24-1.pdf}

@@
